---
output: html_document
editor_options: 
  chunk_output_type: inline
---
# Titratable Acidity, pH, and Sourness {#TA}

```{r library, message=FALSE, warning=FALSE, include=FALSE}
# data wrangling
library(plyr)
library(tidyverse) # data wrangling
library(str2str) # R viz wrangling
library(readxl) # read in data
library(Rmisc) # summarySE()
library(agricolae) #LSD test
library(janitor) #clean_names()
library(tibble)

# correlations
library(Hmisc)

# analysis
library(sensR)

#visualizations
library(ggplot2) # plotting
library(grid) # adjusting axes on plots
library(ggsignif) #add signif to plots
library(cowplot) #combining plots
library(patchwork)
library(gghighlight)
library(knitr)
library(ggpmisc)
library(ggrepel)
```

## Titratable Acidity
```{r ta data, include=FALSE}
TAReadings <- read_excel("~/Research/GitHub/Objective1/TAReadings.xlsx")
TAReadings$Rep <- factor(TAReadings$Rep, levels = c(1, 2, 3))

TA5.17 <- TAReadings %>% filter(Sample %in% c("Whole17-5", "Skim17-5", "Water17-5"))
TA5.8 <- TAReadings %>% filter(Sample %in% c("Whole8-5", "Skim8-5", "Water8-5"))
TA10.17 <- TAReadings %>% filter(Sample %in% c("Whole17-10", "Skim17-10", "Water17-10"))
TA10.8 <- TAReadings %>% filter(Sample %in% c("Whole8-10", "Skim8-10", "Water8-10"))
```

```{r ta wrangling, include=TRUE, comment=""}
aov10.8 <- aov(mL_NaOH ~ Sample + Rep, data = TA10.8)
summary(aov10.8)
print(LSD.test(aov10.8, trt = "Sample")$groups)
# marginal change in TA

aov10.17 <- aov(mL_NaOH ~ Sample + Rep, data = TA10.17)
summary(aov10.17)
print(LSD.test(aov10.17, trt = "Sample")$groups)
# water significantly less acidic than milk additions

aov5.8 <- aov(mL_NaOH ~ Sample + Rep, data = TA5.8)
summary(aov5.8)
print(LSD.test(aov5.8, trt = "Sample")$groups)
# no difference in TA

aov5.17 <- aov(mL_NaOH ~ Sample + Rep, data = TA5.17)
summary(aov5.17)
print(LSD.test(aov5.17, trt = "Sample")$groups)
# water significantly less acidic
```

- AOV model includes Sample and Rep main effects on mL NaOH readings  
  + no rep effects seen in anovas (p > 0.05)
  + at 1:10 and 8% addition, water was significantly different from whole milk addition
  + at 17% addition (both 1:10 and 1:5), water was significantly different from either milk addition

```{r}
coffeelevels <- c("Water", "Skim", "Whole",
            "Black5", "Black10",
            "Skim17-5", "Whole17-5", "Water17-5",
            "Skim8-5", "Whole8-5","Water8-5", 
            "Skim17-10", "Whole17-10", "Water17-10",
            "Skim8-10", "Whole8-10", "Water8-10") 

TASummary <- summarySE(data = TAReadings, measurevar = "mL_NaOH", groupvars = "Sample") %>% select(Sample, mL_NaOH, sd) %>% arrange(factor(Sample, levels = coffeelevels)) %>% slice(-(1:5))

TASummary$groups <- c("A", "A", "B", "A", "A", "A",
                      "A", "A", "B", "AB", "A", "B")

coffeeorder <- c("Skim8-10", "Whole8-10", "Water8-10",
                  "Skim17-10", "Whole17-10", "Water17-10",
                  "Skim8-5", "Whole8-5","Water8-5",
                  "Skim17-5", "Whole17-5", "Water17-5")

TASummary$Sample <- factor(TASummary$Sample, levels = coffeeorder)
``` 


```{r ta_viz, echo=FALSE}
TA_viz <- TASummary %>% 
  ggplot(aes(x = factor(Sample, levels = coffeeorder), y = mL_NaOH)) +
  geom_errorbar(data = TASummary,
                aes(x = Sample, ymin = mL_NaOH-sd, ymax = mL_NaOH+sd), width = 0.25, color = "grey") +
  geom_point() +  
  geom_text(data = TASummary,
            aes(y = 5.5, x = Sample,
            label = groups), color = "black") +   
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  labs(x = "Sample", y = "mL NaOH") +
  scale_y_continuous(expand = c(0,0), limits = c(0, 6)) +
  geom_vline(xintercept = 3.5) +
  geom_vline(xintercept = 6.5) +
  geom_vline(xintercept = 9.5)
```
```{r TAGroupsViz, echo=FALSE}
#| fig.cap = "Titratable acidity (mean +/- standard deviation) of cold brew coffee samples with either skim milk, whole milk, or water added. For example, Skim8-10 has 8% skim milk added to a cold brew coffee at a 1:10 brew strength. Samples within a section that do not share a letter are significantly different as determined by a 2-way ANOVA (sample and rep) and LSD post-hoc test."
TA_viz
```

## TA Percent Acid

$$
\%\ acid\ (w/v) = \frac{N \cdot V_1 \cdot Eq wt}{V_2 \cdot 1000} \cdot 100
$$

where  

- N = normality of titrant
- $V_1$ = volume of titrant
- Eqwt = equivalent weight of relevant acid
+ Eqwt CGA = 354.31
- $V_2$ = volume of solution

```{r percent acid calc, include=FALSE}
head(TAReadings)

N <- 0.1
EqW <- 354.31
V2 <- 25

TAReadings$percent_CGA <- (N*TAReadings$mL_NaOH*EqW)/(V2*1000)*100

TA5.17 <- TAReadings %>% filter(Sample %in% c("Whole17-5", "Skim17-5", "Water17-5"))
TA5.8 <- TAReadings %>% filter(Sample %in% c("Whole8-5", "Skim8-5", "Water8-5"))
TA10.17 <- TAReadings %>% filter(Sample %in% c("Whole17-10", "Skim17-10", "Water17-10"))
TA10.8 <- TAReadings %>% filter(Sample %in% c("Whole8-10", "Skim8-10", "Water8-10"))

```

```{r acid aov, include=FALSE}
aov10.8 <- aov(percent_CGA ~ Sample + Rep, data = TA10.8)
summary(aov10.8)
print(LSD.test(aov10.8, trt = "Sample")$groups)
# marginal difference in %CGA

aov10.17 <- aov(percent_CGA ~ Sample + Rep, data = TA10.17)
summary(aov10.17)
print(LSD.test(aov10.17, trt = "Sample")$groups)
# water significantly less %CGA than milk additions

aov5.8 <- aov(percent_CGA ~ Sample + Rep, data = TA5.8)
summary(aov5.8)
print(LSD.test(aov5.8, trt = "Sample")$groups)
# no difference in TA

aov5.17 <- aov(percent_CGA ~ Sample + Rep, data = TA5.17)
summary(aov5.17)
print(LSD.test(aov5.17, trt = "Sample")$groups)
# water significantly less acidic
```

- AOV model includes Sample and Rep main effects on percent CGA calculations  
  + no rep effects seen in anovas (p > 0.05)
  + at 1:10 and 8% addition, water was significantly different from whole milk addition
  + at 17% addition (both 1:10 and 1:5), water was significantly different from either milk addition



```{r acid_AOV, echo=FALSE}
TASummary
# same groups as mL NaOH
```

```{r acid_viz, echo=FALSE}
TAReadings %>% 
  filter(Sample %in% coffeeorder) %>% 
  ggplot(aes(x = factor(Sample, levels = coffeeorder), y = percent_CGA)) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.25, color = "grey") +
  stat_summary(fun = mean, geom = "point", shape = 19, color = "black") +
  geom_text(data = TASummary,
            aes(y = 1, x = Sample,
            label = groups), color = "black") +   
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  labs(x = "Sample", y = "% CGA (w/v)") +
  scale_y_continuous(expand = c(0,0), limits = c(0, 1.1)) +
  geom_vline(xintercept = 3.5) +
  geom_vline(xintercept = 6.5) +
  geom_vline(xintercept = 9.5)
```

```{r CGAGroupsViz, echo=FALSE}
#| fig.cap = "Percent chlorogenic acid (CGA) (mean +/- standard deviation) of cold brew coffee samples with either skim milk, whole milk, or water added. For example, Skim8-10 has 8% skim milk added to a cold brew coffee at a 1:10 brew strength. Samples within a section that do not share a letter are significantly different as determined by a 2-way ANOVA (sample and rep) and LSD post-hoc test."
TA_viz
```

## Correlation with Sour Ratings
### TA Readings + Sourness
```{r import sour ratings, include=FALSE}
# 1:10 Nose Closed
brew10closedsour <- read_xlsx(path = "Objective1CompiledDataWithIndex.xlsx", sheet = 3) 
brew10closedsour <- t(brew10closedsour) %>% as.data.frame() %>% row_to_names(row_number = 1) %>% clean_names()
chars <- sapply(brew10closedsour, is.character)
brew10closedsour[ ,chars] <- as.data.frame(apply(brew10closedsour[,chars],2, as.numeric))
brew10closedsour <- brew10closedsour %>% rownames_to_column(var = "participant")
brew10closedsour_ratings <- brew10closedsour %>% select(participant, starts_with("sourness"))
brew10closedsour_ratings_long <- brew10closedsour_ratings %>% 
  pivot_longer(-participant, names_to = "Sample", values_to = "Intensity")
sample_order10 <- c("sourness_13", "sourness_15", "sourness_14", "sourness_17", "sourness_16", "sourness_18", "sourness_19", "sourness_21", "sourness_20", "sourness_23", "sourness_22", "sourness_24")
brew10closedsour_ratings_long$coffee <- rep("10", length(brew10closedsour_ratings_long$participant))
brew10closedsour_ratings_long$cond <- rep("closed", length(brew10closedsour_ratings_long$participant))


# 1:10 Nose Opened
brew10openedSour <- read_xlsx(path = "Objective1CompiledDataWithIndex.xlsx", sheet = 12)
df10openedSour <- t(brew10openedSour) %>% as.data.frame() %>% row_to_names(row_number = 1) %>% clean_names()
chars <- sapply(df10openedSour, is.character)
df10openedSour[ ,chars] <- as.data.frame(apply(df10openedSour[,chars],2, as.numeric))
df10openedSour <- df10openedSour %>% rownames_to_column(var = "participant")
df10openedSour_ratings <- df10openedSour %>% select(participant, starts_with("sourness"))
df10openedSour_ratings_long <- df10openedSour_ratings %>% 
  pivot_longer(-participant, names_to = "Sample", values_to = "Intensity")
df10openedSour_ratings_long$coffee <- rep("10", length(df10openedSour_ratings_long$participant))
df10openedSour_ratings_long$cond <- rep("opened", length(df10openedSour_ratings_long$participant))

# 1:5 Nose Closed
brew5closedsour <- read_xlsx(path = "Objective1CompiledDataWithIndex.xlsx", sheet = 7) 
brew5closedsour <- t(brew5closedsour) %>% as.data.frame() %>% row_to_names(row_number = 1) %>% clean_names()
chars <- sapply(brew5closedsour, is.character)
brew5closedsour[ ,chars] <- as.data.frame(apply(brew5closedsour[,chars],2, as.numeric))
brew5closedsour <- brew5closedsour %>% rownames_to_column(var = "participant")
brew5closedsour_ratings <- brew5closedsour %>% select(participant, starts_with("sourness"))
brew5closedsour_ratings_long <- brew5closedsour_ratings %>% 
  pivot_longer(-participant, names_to = "Sample", values_to = "Intensity")
sample_order5 <- c("sourness_1", "sourness_3", "sourness_2", "sourness_5", "sourness_4", "sourness_6", "sourness_7", "sourness_9", "sourness_8", "sourness_11", "sourness_10", "sourness_12")
brew5closedsour_ratings_long$coffee <- rep("5", length(brew5closedsour_ratings_long$participant))
brew5closedsour_ratings_long$cond <- rep("closed", length(brew5closedsour_ratings_long$participant))

# 1:5 Nose Opened
brew5openedSour <- read_xlsx(path = "Objective1CompiledDataWithIndex.xlsx", sheet = 11)
df5openedSour <- t(brew5openedSour) %>% as.data.frame() %>% row_to_names(row_number = 1) %>% clean_names()
chars <- sapply(df5openedSour, is.character)
df5openedSour[ ,chars] <- as.data.frame(apply(df5openedSour[,chars],2, as.numeric))
df5openedSour <- df5openedSour %>% rownames_to_column(var = "participant")
df5openedSour_ratings <- df5openedSour %>% select(participant, starts_with("sourness"))
df5openedSour_ratings_long <- df5openedSour_ratings %>% 
  pivot_longer(-participant, names_to = "Sample", values_to = "Intensity")
df5openedSour_ratings_long$coffee <- rep("5", length(df5openedSour_ratings_long$participant))
df5openedSour_ratings_long$cond <- rep("opened", length(df5openedSour_ratings_long$participant))
```

```{r rename samples}
unique(brew10closedsour_ratings_long$Sample)
brew10closedsour_ratings_long <- brew10closedsour_ratings_long %>% 
  mutate(Sample = recode(Sample, "sourness_13" = "Skim17-10")) %>% 
  mutate(Sample = recode(Sample, "sourness_14" = "Skim17-10")) %>% 
  mutate(Sample = recode(Sample, "sourness_15" = "Whole17-10")) %>% 
  mutate(Sample = recode(Sample, "sourness_16" = "Whole17-10")) %>% 
  mutate(Sample = recode(Sample, "sourness_17" = "Water17-10")) %>% 
  mutate(Sample = recode(Sample, "sourness_18" = "Water17-10")) %>% 
  mutate(Sample = recode(Sample, "sourness_19" = "Skim8-10")) %>% 
  mutate(Sample = recode(Sample, "sourness_20" = "Skim8-10")) %>% 
  mutate(Sample = recode(Sample, "sourness_21" = "Whole8-10")) %>% 
  mutate(Sample = recode(Sample, "sourness_22" = "Whole8-10")) %>% 
  mutate(Sample = recode(Sample, "sourness_23" = "Water8-10")) %>% 
  mutate(Sample = recode(Sample, "sourness_24" = "Water8-10")) 

unique(df10openedSour_ratings_long$Sample)
df10openedSour_ratings_long <- df10openedSour_ratings_long %>% 
  mutate(Sample = recode(Sample, "sourness_13" = "Skim17-10")) %>% 
  mutate(Sample = recode(Sample, "sourness_14" = "Skim17-10")) %>% 
  mutate(Sample = recode(Sample, "sourness_15" = "Whole17-10")) %>% 
  mutate(Sample = recode(Sample, "sourness_16" = "Whole17-10")) %>% 
  mutate(Sample = recode(Sample, "sourness_17" = "Water17-10")) %>% 
  mutate(Sample = recode(Sample, "sourness_18" = "Water17-10")) %>% 
  mutate(Sample = recode(Sample, "sourness_19" = "Skim8-10")) %>% 
  mutate(Sample = recode(Sample, "sourness_20" = "Skim8-10")) %>% 
  mutate(Sample = recode(Sample, "sourness_21" = "Whole8-10")) %>% 
  mutate(Sample = recode(Sample, "sourness_22" = "Whole8-10")) %>% 
  mutate(Sample = recode(Sample, "sourness_23" = "Water8-10")) %>% 
  mutate(Sample = recode(Sample, "sourness_24" = "Water8-10")) 

unique(brew5closedsour_ratings_long$Sample)
brew5closedsour_ratings_long <- brew5closedsour_ratings_long %>% 
  mutate(Sample = recode(Sample, "sourness_1" = "Skim17-5")) %>% 
  mutate(Sample = recode(Sample, "sourness_2" = "Skim17-5")) %>% 
  mutate(Sample = recode(Sample, "sourness_3" = "Whole17-5")) %>% 
  mutate(Sample = recode(Sample, "sourness_4" = "Whole17-5")) %>% 
  mutate(Sample = recode(Sample, "sourness_5" = "Water17-5")) %>% 
  mutate(Sample = recode(Sample, "sourness_6" = "Water17-5")) %>% 
  mutate(Sample = recode(Sample, "sourness_7" = "Skim8-5")) %>% 
  mutate(Sample = recode(Sample, "sourness_8" = "Skim8-5")) %>% 
  mutate(Sample = recode(Sample, "sourness_9" = "Whole8-5")) %>% 
  mutate(Sample = recode(Sample, "sourness_10" = "Whole8-5")) %>% 
  mutate(Sample = recode(Sample, "sourness_11" = "Water8-5")) %>% 
  mutate(Sample = recode(Sample, "sourness_12" = "Water8-5")) 

unique(df5openedSour_ratings_long$Sample)
df5openedSour_ratings_long <- df5openedSour_ratings_long %>% 
  mutate(Sample = recode(Sample, "sourness_1" = "Skim17-5")) %>% 
  mutate(Sample = recode(Sample, "sourness_2" = "Skim17-5")) %>% 
  mutate(Sample = recode(Sample, "sourness_3" = "Whole17-5")) %>% 
  mutate(Sample = recode(Sample, "sourness_4" = "Whole17-5")) %>% 
  mutate(Sample = recode(Sample, "sourness_5" = "Water17-5")) %>% 
  mutate(Sample = recode(Sample, "sourness_6" = "Water17-5")) %>% 
  mutate(Sample = recode(Sample, "sourness_7" = "Skim8-5")) %>% 
  mutate(Sample = recode(Sample, "sourness_8" = "Skim8-5")) %>% 
  mutate(Sample = recode(Sample, "sourness_9" = "Whole8-5")) %>% 
  mutate(Sample = recode(Sample, "sourness_10" = "Whole8-5")) %>% 
  mutate(Sample = recode(Sample, "sourness_11" = "Water8-5")) %>% 
  mutate(Sample = recode(Sample, "sourness_12" = "Water8-5")) 

colnames(df5openedSour_ratings_long)
brew10sour <- full_join(brew10closedsour_ratings_long, df10openedSour_ratings_long, by = c("participant", "Sample", "Intensity", "coffee", "cond"))
brew5sour <- full_join(brew5closedsour_ratings_long, df5openedSour_ratings_long, by = c("participant", "Sample", "Intensity", "coffee", "cond"))
sourDF <- full_join(brew10sour, brew5sour)
view(sourDF)

sourTAjoined <- full_join(sourDF, TASummary)
view(sourTAjoined)
```



```{r sour rcorr closed, echo=TRUE, include=FALSE, eval=FALSE}
sourSkim17.5C <- sourTAjoined %>% filter(Sample == "Skim17-5" & cond == "closed") %>% select(mL_NaOH, Intensity) %>% as.matrix()
(sourSkim17.5C <- rcorr(sourSkim17.5C, type = "pearson")) #p < 0.001
sourSkim17.5O <- sourTAjoined %>% filter(Sample == "Skim17-5" & cond == "opened") %>% select(mL_NaOH, Intensity) %>% as.matrix()
(sourSkim17.5O <- rcorr(sour_to_cor_closed, type = "pearson")) #p < 0.001
sourWhole17.5C <- sourTAjoined %>% filter(Sample == "Whole17-5" & cond == "closed") %>% select(mL_NaOH, Intensity) %>% as.matrix()
(sourWhole17.5C <- rcorr(sour_to_cor_closed, type = "pearson")) #p < 0.001
sourWhole17.5O <- sourTAjoined %>% filter(Sample == "Whole17-5" & cond == "opened") %>% select(mL_NaOH, Intensity) %>% as.matrix()
(sourWhole17.5O <- rcorr(sour_to_cor_closed, type = "pearson")) #p < 0.001
sourWater17.5C <- sourTAjoined %>% filter(Sample == "Water17-5" & cond == "closed") %>% select(mL_NaOH, Intensity) %>% as.matrix()
(sourWater17.5C <- rcorr(sour_to_cor_closed, type = "pearson")) #p < 0.001
sourWater17.5O <- sourTAjoined %>% filter(Sample == "Water17-5" & cond == "opened") %>% select(mL_NaOH, Intensity) %>% as.matrix()
(sourWater17.5O <- rcorr(sour_to_cor_closed, type = "pearson")) #p < 0.001
##
sourSkim8.5C <- sourTAjoined %>% filter(Sample == "Skim8-5" & cond == "closed") %>% select(mL_NaOH, Intensity) %>% as.matrix()
(sourSkim8.5C <- rcorr(sour_to_cor_closed, type = "pearson")) #p < 0.001
sourSkim8.5O <- sourTAjoined %>% filter(Sample == "Skim8-5" & cond == "opened") %>% select(mL_NaOH, Intensity) %>% as.matrix()
(sourSkim8.5O <- rcorr(sour_to_cor_closed, type = "pearson")) #p < 0.001
sourWhole8.5C <- sourTAjoined %>% filter(Sample == "Whole8-5" & cond == "closed") %>% select(mL_NaOH, Intensity) %>% as.matrix()
(sourWhole8.5C <- rcorr(sour_to_cor_closed, type = "pearson")) #p < 0.001
sourWhole8.5O <- sourTAjoined %>% filter(Sample == "Whole8-5" & cond == "opened") %>% select(mL_NaOH, Intensity) %>% as.matrix()
(sourWhole8.5O <- rcorr(sour_to_cor_closed, type = "pearson")) #p < 0.001
sourWater8.5C <- sourTAjoined %>% filter(Sample == "Water8-5" & cond == "closed") %>% select(mL_NaOH, Intensity) %>% as.matrix()
(sourWater8.5C <- rcorr(sour_to_cor_closed, type = "pearson")) #p < 0.001
sourWater8.5O <- sourTAjoined %>% filter(Sample == "Water8-5" & cond == "opened") %>% select(mL_NaOH, Intensity) %>% as.matrix()
(sourWater8.5O <- rcorr(sour_to_cor_closed, type = "pearson")) #p < 0.001
##
sourSkim17.10C <- sourTAjoined %>% filter(Sample == "Skim17-10" & cond == "closed") %>% select(mL_NaOH, Intensity) %>% as.matrix()
(sourSkim17.10C <- rcorr(sour_to_cor_closed, type = "pearson")) #p < 0.001
sourSkim17.10O <- sourTAjoined %>% filter(Sample == "Skim17-10" & cond == "opened") %>% select(mL_NaOH, Intensity) %>% as.matrix()
(sourSkim17.10O <- rcorr(sour_to_cor_closed, type = "pearson")) #p < 0.001
sourWhole17.10C <- sourTAjoined %>% filter(Sample == "Whole17-10" & cond == "closed") %>% select(mL_NaOH, Intensity) %>% as.matrix()
(sourWhole17.10C <- rcorr(sour_to_cor_closed, type = "pearson")) #p < 0.001
sourWhole17.10O <- sourTAjoined %>% filter(Sample == "Whole17-10" & cond == "opened") %>% select(mL_NaOH, Intensity) %>% as.matrix()
(sourWhole17.10O <- rcorr(sour_to_cor_closed, type = "pearson")) #p < 0.001
sourWater17.10C <- sourTAjoined %>% filter(Sample == "Water17-10" & cond == "closed") %>% select(mL_NaOH, Intensity) %>% as.matrix()
(sourWater17.10C <- rcorr(sour_to_cor_closed, type = "pearson")) #p < 0.001
sourWater17.10O <- sourTAjoined %>% filter(Sample == "Water17-10" & cond == "opened") %>% select(mL_NaOH, Intensity) %>% as.matrix()
(sourWater17.10O <- rcorr(sour_to_cor_closed, type = "pearson")) #p < 0.001
##
sourSkim8.10C <- sourTAjoined %>% filter(Sample == "Skim8-10" & cond == "closed") %>% select(mL_NaOH, Intensity) %>% as.matrix()
(sourSkim8.10C <- rcorr(sour_to_cor_closed, type = "pearson")) #p < 0.001
sourSkim8.10O <- sourTAjoined %>% filter(Sample == "Skim8-10" & cond == "opened") %>% select(mL_NaOH, Intensity) %>% as.matrix()
(sourSkim8.10O <- rcorr(sour_to_cor_closed, type = "pearson")) #p < 0.001
sourWhole8.10C <- sourTAjoined %>% filter(Sample == "Whole8-10" & cond == "closed") %>% select(mL_NaOH, Intensity) %>% as.matrix()
(sourWhole8.10C <- rcorr(sour_to_cor_closed, type = "pearson")) #p < 0.001
sourWhole8.10O <- sourTAjoined %>% filter(Sample == "Whole8-10" & cond == "opened") %>% select(mL_NaOH, Intensity) %>% as.matrix()
(sourWhole8.10O <- rcorr(sour_to_cor_closed, type = "pearson")) #p < 0.001
sourWater8.10C <- sourTAjoined %>% filter(Sample == "Water8-10" & cond == "closed") %>% select(mL_NaOH, Intensity) %>% as.matrix()
(sourWater8.10C <- rcorr(sour_to_cor_closed, type = "pearson")) #p < 0.001
sourWater8.10O <- sourTAjoined %>% filter(Sample == "Water8-10" & cond == "opened") %>% select(mL_NaOH, Intensity) %>% as.matrix()
(sourWater8.10O <- rcorr(sour_to_cor_closed, type = "pearson")) #p < 0.001
```
```{r}
sourTAjoined %>% drop_na() %>% 
  ggplot(aes(x = mL_NaOH, y = Intensity)) +
  facet_grid(coffee~cond) +
  geom_boxplot(aes(fill = Sample), color = "black") +
  stat_summary(fun = mean, geom = "point", shape = 21, size = 1) +
  theme_bw() +
  scale_y_continuous(expand = c(0,.2))
```

